### 2014-12-26 mcmc.tex
methods content

% style notes
% -----------
% - [LaTeX] newline after every full stop for proper git diffing
% - [LaTeX] \eqnarray not \equation
% - pdf not PDF
% - is it ``a MCMC sampling'' or ``an MCMC sampling''?
% - \note{} right after word if you are endnoting a phrase,
%   but after the full stop if endnoting the sentence or paragraph.
% - do we work in $x$ or $\pars$?  I am confused...

@@

\thispagestyle{plain}\raggedbottom
\section*{Data analysis recipes:\\ \this\footnotemark}

@@

\section{Gibbs, Ensembles, Hamiltonians, and other variants}

@@

...What, briefly, is the stretch move, and why is it useful?

>>>>

% style notes
% -----------
% - [LaTeX] newline after every full stop for proper git diffing
% - [LaTeX] \eqnarray not \equation
% - pdf not PDF
% - is it ``a MCMC sampling'' or ``an MCMC sampling''?
% - \note{} right after word if you are endnoting a phrase,
%   but after the full stop if endnoting the sentence or paragraph.
% - do we work in $x$ or $\pars$?  I am confused...
% - should we use the word ``walker'' anywhere, everywhere, or never?

@@

\thispagestyle{plain}\sloppy\sloppypar\raggedbottom
\section*{Data analysis recipes:\\ \this\footnotemark}

@@

\section{Gibbs, ensemble, Hamiltonian, and so on}

MCMC methods are proven to be correct ``in the limit''; that is, when
an infinite number of samples have been taken.
When the function $f(\pars)$ has challenging properties, the results
can approach this limit very slowly (or not at all\note{HOGG: discuss
  situations in which MCMC will not work---related to support.}).
For this reason, in the finite duration of a scientific project, most
users of MCMC have one of two (related) problems:
Either it is taking too long---they have take too many steps or
execute too many calls of the function $f(\pars)$---or it is not
exploring the full parameter space---there are local optima (or bad
local geometry around those optima) ``trapping'' the algorithm.

There is no general, problem-independent solution for either of these
problems.
Indeed, there are many different strategies that work well or badly
for different functions $f(\pars)$.
In this \sectionname, we will try to guide the reader towards methods
that might help in different circumstances.
We will not give a full description of all the MCMC algorithms out
there, but just provide some field notes and pointers to references.
New MCMC methods are being developed all the time, so a reader with a
serious problem would do well to consult with the applied-mathematics
or statistics literature.\note{HOGG: Your best interface to these
  literatures is via a colleague.  Break down those walls!}
The different methods to which we refer here are (in general) very
different in their difficulty of implementation.
We won't help with that either, except to say that there are more and
more open-source or distributed packages every year.
We will try to name some current implementations in the notes, but any
list compiled today will be incomplete tomorrow.
The spaces of MCMC methods and MCMC method implementations are both
growing rapidly, in part because MCMC has become such a core
technology in the empirical sciences.

@@

...Gibbs!  Every parameter is differently difficult to sample in?

...What, briefly, is the stretch move, and why is it useful?  Every parameter treated identically---almost opposite to gibbs.
